---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
The zip of the data  is present on the root directory
Unzip it and read the data
``` {r Unzipandread, echo=TRUE}
unzip("activity.zip")
data<-read.csv("activity.csv")
```

Have a quick look at the data using summary
```{r Summary, echo=TRUE}
summary(data)
```

And the str
```{r Str, echo=TRUE}
str(data)
```
Looks good like : keep the data like that

## What is mean total number of steps taken per day?

In this section we keept the missing values.
Aggregate the the steps versus date and sum it to have a set of data with just the total number of steps
```{r TotalStep, echo=TRUE}
totalSteps<-aggregate(steps~date,data=data,sum,na.rm=TRUE)
head(totalSteps,5)
```

Make an histogram of the total number of steps per day
```{r Histo, echo=TRUE}
hist(totalSteps$steps, xlab="Total Steps",  main =" Histogram of total number of steps per day")
```

Calculate the mean and the median value of the total number of steps.
```{r MeanandMedian, echo=TRUE}
mean(totalSteps$steps)
median(totalSteps$steps)
```


## What is the average daily activity pattern?
Caculate the mean steps of daily activity
```{r CalAveActivity, echo=TRUE}
AverageActivity<-aggregate(steps~interval,data=data,mean,na.rm=TRUE)
head(AverageActivity,5)
```

Plot the pattern found
```{r PlotAv, echo=TRUE}
plot(steps~interval,data=AverageActivity,type="l",main="Average Activity Pattern",xlab="activity by 5mn",ylab="Average steps")
```

Retrieve the interval which contain the max of mean step.
```{r MaxInterval, echo=TRUE}
Interval_MaxMeanSteps = AverageActivity[which.max(AverageActivity[, 2]), 1]
Interval_MaxMeanSteps
```

## Imputing missing values

there are a number of days/intervals where there are missing values (coded as ð™½ï¿½A).

The presence of missing days may introduce bias into some calculations or summaries of the data.

calculate the number of rows with NA
```{r calculNATot, echo=TRUE}
sum(is.na(data$steps))
```
strategy to fill in NA value: 

create a new data set (data2) set with the NA replace by mean of t 5 mn (already computed in the previous question)
```{r Strategy, echo=TRUE}
Newdata = data
for (i in 1:nrow(Newdata)) {
  if (is.na(Newdata$steps[i])) {
    for (j in 1:nrow(AverageActivity)) {
      if (Newdata$interval[i] == AverageActivity[j, 1]) {
        Newdata$steps[i] = AverageActivity[j, 2]
      }
    } 
  }    
}
```
Verify the set is not conatining NA now
```{r verification new set, echo=TRUE}
sum(is.na(Newdata$steps))
head(Newdata,5)
```

Recalculate the total steps by day, plot histogram, calculate mean and median
```{r PlotasFirst, echo=TRUE}

NewtotalSteps<-aggregate(steps~date,data=Newdata,sum,na.rm=TRUE)

hist(NewtotalSteps$steps, xlab="Total Steps",  main =" Histogram of total number of steps per day")

mean(NewtotalSteps$steps)
median(NewtotalSteps$steps)
```
We obtained an histogram nearly smilar to the previous ( some value change but not the form of it)

we obtained the same mean as previous but the median value is now egal to the mean. This impact is  due by the replacement of the 2304 missing value by the 5-min interval obtained from the original dataset.

## Are there differences in activity patterns between weekdays and weekends?

fisrt we have to add a new column in the previous dataset (name day).

(local set to have english name of day to be reproduced by anyone.
```{r Newcolumn, echo=TRUE}
Sys.setlocale("LC_TIME", "C")
#Sys.setenv(LANG="en")
Newdata$day_of_week <- weekdays(as.Date(Newdata$date), abbreviate = TRUE)
head(Newdata) 
```

Assiociate day with day type (weekday or weekend) and add it to the dataset
```{r fillit, echo=TRUE}
days <- c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")
day_types <- c(rep("weekday", 5), rep("weekend", 2))
days_of_week <- data.frame(days, day_types, stringsAsFactors = FALSE)
colnames(days_of_week) <- c("day_of_week", "day_type")
Newdata <- merge(Newdata, days_of_week, by = "day_of_week")
Newdata$day_type <- as.factor(Newdata$day_type)
head(Newdata)
```

use lattice for a similar look as the example

recalculate the daily activity pattern 

Check the average data frame

plot the two figure for weekdays and weekend
```{r  2plots, echo=TRUE}
library(lattice)

AverageActivity2=aggregate(steps~interval+day_type,Newdata,mean)
head(AverageActivity2,3)
xyplot(steps~interval|factor(day_types),data=AverageActivity,aspect=1/2,type="l",main="Average Activity Pattern",xlab="activity by 5mn",ylab="Average steps")
```